---
title: "Serial reading with Python; data processing with R"
date: '2021-04-22'
output:
  powerpoint_presentation:
    reference_doc: serial-reading-template.pptx
---

```{r setup, include=F}
knitr::opts_chunk$set(comment = "", fig.retina = 3)
```

# Writing to a file via USB 

- Not as simple as it would seem 

- Once working, minimizes data handling pipeline

# Python solution

- `pyserial` module and `csv` module 

- `while` loop to maintain stream 


# Python code 

````python
import serial
import time
import csv
````
````python
ser = serial.Serial('COM4', 115200)
ser.flushInput()
````
````python
while True:
    try:
        ser_bytes = ser.readline()
       
        print(ser_bytes)
        with open("test_data.csv","a") as f:
            writer = csv.writer(f,delimiter=",")
            writer.writerow([time.time(), ser_bytes])
    except:
        print("Keyboard Interrupt")
        break
````

# Clean up with **tidyverse**

```{r}
# Read the csv file and specify column names since no header is present 

library(magrittr)

df <- readr::read_delim(file = './test_data.csv',
                 col_names = c("read_time", "message"),
                 delim = ",",
                 col_types = 'dc')
```

# Clean up with **tidyverse**

```{r}
head(df, n = 10)
```

---

# Clean up with **tidyverse**

```{r}
# parse and clean data 
# employs a couple beefy regular expressions 
df <- df %>% 
  dplyr::mutate(
    read_time = lubridate::as_datetime(read_time),
    message = stringr::str_remove_all(message, pattern = "b'|\\\\r|\\\\n'")) %>%
  dplyr::filter(
    stringr::str_detect(message, "\\w{4}\\s\\w{4}\\s\\w{4}\\s\\w{4}" )) %>% 
  dplyr::mutate(
    message = stringr::str_remove(message, "\\s*UID\\sValue:\\s")) %>% 
  dplyr::rename(card_id = message)
```

# Clean up with **tidyverse**

:::::: {.columns}

::: {.column}
- `datetime` objects represent the number of seconds since beginning of Unix epoch (Jan 1, 1970 00:00 UTC)  


- They are based on a double-type atomic vector:
```{r}
unclass(df$read_time)[1]
```


:::

::: {.column}

- exist in both Python and R

- detection time limited to 1-s precision w/ Python `time.time()`


:::

::::::
---


```{r time-series, fig.asp=9/16, out.width='100%', echo=F}
library(ggplot2)

ggplot(df, aes(read_time, 1))+
  geom_col(fill = 'firebrick', alpha = 1/4)+
  scale_x_datetime('Detection time',
                   labels = scales::label_date(format = "%r", tz = "America/New_York"), expand = expansion(mult = 0.12))+
  theme_void()+
  labs(title = 'Detection times of NFC RFID card')+
  theme(
    axis.line.x = element_line(),
    axis.ticks.x = element_line(),
    axis.text.x = element_text(hjust = 1)
  )
  
  
  

```

